### 2023-10-20 README.Rmd
updated tests and formatted function - 231020

---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# validateHOT 🎯

<!-- badges: start -->

[![R-CMD-check](https://github.com/JoshSchramm94/validateHOT/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JoshSchramm94/validateHOT/actions/workflows/R-CMD-check.yaml)

[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
<!-- badges: end -->

The goal of validateHOT is to validate the results of your validation
task (also known as holdout task). A validation task is essential to
make sure that your collected data of a MaxDiff, CBC, or ACBC is valid
and can also predict outside tasks that were not included in estimating
your utility scores. Although commercial studies often do not include a
validation/holdout task (Yang et al., 2018), it is highly recommended to do so
(Orme, 2015; Rao, 2014). This validation/holdout task does not only
help to check whether everything went right during data collection but
also to determine your final model. <code>validatHOT</code> provides
some of the relevant metrics to test the performance of your data in
predicting a holdout task. In [Sawtooth
Software's](https://sawtoothsoftware.com/) CBC a fixed validation/
holdout task is automatically implemented. If you conduct a MaxDiff or
ACBC study, you have to program these on your own.

> Currently <code>validateHOT</code> only works with data from Sawtooth Software 
due to the coding of price.

👉🏾 <u>**What you need to provide**</u>: <br> After collecting your data,
and running your initial Hierarchical Bayes models, you can turn to
<code>validateHOT</code> and test how good your model predicts choices
in the validation/holdout task. Herefore, you only have to read in your
raw utility scores as well as the actual choice of your validation/
holdout task. You can use the <code>merge()</code> provided by the
<code>base</code> package (R Core Team, 2023). Afterward, you can read in your data
file and enjoy <code>validateHOT</code>. We provide a tutorial a short tutorial in the vignette.

👈🏾 <u>**What you get**</u>:<br> At the moment, <code>validateHOT</code>
provides functions for 3 key areas:

<ul>

<li>validation metrics mainly reported in preference measurement
studies</li>

<li>metrics that are usually reported in machine learning (i.e.,
confusion matrix)</li>

<li>simulation methods, for example, to determine optimal product
combinations</li>

</ul>

For all 3 key areas, the <code>createHOT</code> function is essential.
This function creates the total utilities for each alternative in the
validation/holdout task and also in the simulation method, respectively.
Rao (2014, p. 82) mentions the additive utility model stating that the
total utility of a profile or conjoint model is the sum of its attribute
levels. <code>createHOT</code> will do exactly this for you.

### Classical validation metrics

<ul>

<li><code>hitrate</code>: creates the Hit Rate (correctly predicted
choices) of your validation task. The output will contain 
the percentage of how many choices were correctly predicted, 
the chance level in your validation task ($\frac{1}{alternatives}$) in percentage,
the number of correctly predicted participants' choices as well as the number of observations. 
</li>

<li><code>kl</code>: Kullback-Leibler-Divergence measures the
divergence between the actual choice distribution and the predicted
choice distribution (Ding et al., 2011; Drost, 2018). Output provides both
divergence between predicted from observed and observed from predicted
due to the asymmetry of the Kullback-Leibler divergence. If you specify
an optional <code>group</code> argument the output is split by
groups. Currently, you can choose between \eqn{log} and \eqn{log{_2}} as logarithm base.
The default is set to \eqn{log}.</li>

<li><code>mae</code>: mean absolute error, i.e., deviation between
predicted and stated choice share. </li>

<li><code>medae</code>: median absolute error, which is less affected
by outliers compared to the mean absolute error.</li>

<li><code>mhp</code>: averaged hit probability of participant's actual
choice in the validation/holdout task.</li>

<li><code>rmse</code>: provides the root-mean-square error of deviation
between predicted and stated choice share.</li>

</ul>

All functions can be extended with the `group` argument to get output by group.

### Confusion Matrix

We also include metrics from machine learning, i.e., the confusion
matrix (e.g., Burger, 2018). For all of the 5 provided functions, you
currently have to include a **none** alternative in your validation/holdout task. We currently
predict, e.g., whether a buy or no-buy was correctly predicted.
Information could be used for overestimating and underestimating,
respectively, of product purchases. In the following <code>TP</code>
stands for true positives, <code>FP</code> for false positives,
<code>TN</code> for true negatives, and <code>FN</code> for false
negatives (Burger, 2018). 
To translate this to the `validateHOT` logic, imagine you have a validation/holdout task
with 5 alternatives plus the alternative not to buy any of those chosen. `validateHOT` now measures
whether or not a buy (participant opts for one of the 5 alternatives) or a no-buy (participant opts
for the no-buy alternative), respectively, is correctly predicted.

<ul>

<li><code>accuracy</code>: calculates the number of correctly predicted choices (buy or no-buy);
$\frac{TP + TN}{TP + TN + FP + FN}$ (Burger, 2018). 
</li>

<li><code>f1</code>: defined as
$\frac{2 * precision * recall}{precision + recall}$ or stated
differently by Burger (2018) $\frac{2TP}{2TP + FP + FN}$.</li>

<li><code>precision</code>: defined as $\frac{TP}{TP + FP}$ (Burger, 2018).</li>

<li><code>recall</code>: defined as $\frac{TP}{TP + FN}$ (Burger, 2018).</li>

<li><code>specificity</code>: defined as $\frac{TN}{TN + FP}$ (Burger, 2018).</li>

</ul>

Again, all functions can be extended with the `group` argument to get output by group.

### Simulation Methods

<ul>

<li><code>turf</code>: <b>T</b>(otal) <b>U</b>(nduplicated) <b>R</b>(each) and <b>F</b>(requency) is a "product line extension model" (Miaoulis et al., 1990, p. 29) that helps to find the perfect product bundle based on the <code>reach</code> (e.g., how many participants consider buying at least one product of that assortment) and the <code>frequency</code> (how many products are on average a purchase option). <code>turf</code> currently provides both the threshold approach (<code>approach = 'thres'</code>; all products that exceed a threshold are considered, e.g., a purchase option; Chrzan & Orme, 2019, p. 112) and the first choice approach (<code>approach = 'fc'</code>; only product with highest utility is considered as purchase option; Chrzan & Orme, 2019, p. 111).</li>

<li><code>freqassort</code>: Similar to <code>turf</code>, <code>freqassort</code>
will give you the averaged frequency, how many products the participants
will choose from your in the function determined potential assortment. 
Again, you have to define a <code>none</code> alternative, because <code>freqassort</code> uses the <i>threshold</i> approach,
meaning if the utility of one product is above the utility of <code>none</code>,
it is marked as potential purchase option. While <code>turf</code> calculates the reach and frequency for <b>all</b> combinations, you specify the combination you are interested in <code>freqassort</code>.</li>

<li><code>reach</code>: Similar to <code>turf</code>, <code>reach</code> will
give you the averaged percentage of how many participants you can reach
(at least one of the products resemble a purchase option) with your in the function determined
potential assortment. <code>reach</code> also uses the <i>threshold</i> approach (see above). While <code>turf</code> calculates the reach and frequency for <b>all</b> combinations, you specify the combination you are interested in <code>reach</code></li>

<li><code>marksim</code>: runs market simulations (either share of preference or first choice rule), including the standard error, as well as the lower and upper confidence interval, which is
calculated according to the $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$ (Orme, 2020, p. 94).</li>

</ul>

### Data Frames provided by <code>validateHOT</code>

<ul>

<li><code>ACBC</code>: Example data set with raw utilities of an ACBC
study conducted in Sawtooth. Price was linear-coded while the other
attributes were coded as part-worths.</li>

<li><code>ACBC_interpolate</code>: Example data set with raw utilities
of an ACBC study conducted in Sawtooth. Price was piecewise-coded,
another attribute was linear-coded while the other attributes were coded
as part-worths.</li>

<li><code>CBC</code>: Example data set with raw utilities of an CBC
study conducted in Sawtooth. All attributes were coded as
part-worth.</li>

<li><code>CBC_lin</code>: Example data set with raw utilities of an CBC
study conducted in Sawtooth. One attribute was linear coded while the
other attributes are part-worth coded.</li>

<li><code>MaxDiff</code>: Example data set with raw utilities of an
MaxDiff study conducted in Sawtooth.</li>

</ul>

## Why <code>validateHOT</code>

We are teaching a preference measurement seminar for students. Often
these students did not have any prior experience (or only sparsely) with *R*.
One of the chapters in this class is about model validating the results obtained and we teach this,
of course, in *R*.
Of course, there are other great packages which are faster in running (i.e., `Metrics` by Hamner & Frasco, 2018), 
however, these packages need some more data wrangling in order to use the
appropriate functions, which might be a burden or barrier for one or
the other.

Moreover, as Yang et al. (2018) report, commercial studies often do not use
any validation task. Again, the missing experience in *R* could be one explanation. 
Since these functions are not always implemented in other softwares, this might be one reason why
they do not include one simply because they do not know how to use it
correctly. Having a package to evaluate the validation/holdout task can
also be beneficial from this perspective.

## Installation

You can install the development version of validateHOT from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("JoshSchramm94/validateHOT")
```

## Example

First, we load the package.

```{r}
library("validateHOT")
```

### Example I - CBC

Since *CBC's* are most commonly used between *CBC*, *ACBC*, and
*MaxDiff*, we will provide an example with a *CBC*. Let us begin with a
*CBC* where all of the variables are coded as part-worth, so we need no
interpolation. Let us load the <code>CBC</code> data frame for this
example.

```{r}
data("CBC")
```

The data frame has a total of `r nrow(CBC)` participants and `r ncol(CBC)` columns.

Now imagine you included a validation/ holdout task with three
alternatives as well as an no-buy option. We specify the <code>data</code> argument and the column index of <code>id</code>. Since we also have a *no-buy* alternative in our validation task, we next specify the <code>none</code> argument, otherwise we would have left it empty. Afterwards, we specify the number of products (<code>prod</code>) we include (excluding the *no-buy* alternative), which we define in the next step with the argument <code>prod.levels</code>. If we look back at the data frame, we can see that the first alternative in the holdout task <code>c(4, 9, 19)</code> is composed of the following attribute levels <code>Att1_Lev1</code>, <code>Att2_Lev1</code>, and <code>Att3_Lev5</code>.

As mentioned above, all the attributes are part-worth coded, therefore, we set <code>coding = c(0, 0, 0)</code>. Finally, we specify the method of preference measurement technique, which is <code>method = "CBC"</code> in our case and the column index of the actual participant's choice (<code>choice</code>). If you run the code, a data frame called <code>HOT</code> (short form of **H**old**o**ut **t**ask) will be returned to the global environment.

> &#10071; <code>createHOT</code> is currently just taking indexes instead of column names, please be aware of this. However, you can easily find out the index by using the <code>names()</code> function or by using <code>which()</code> and <code>colnames()</code>, both functions are provided by the base package (R Core Team, 2023). For example, if you want to find out the column index of <code>ID</code>, we could also use <code>which(colnames(CBC) == "ID")</code>.

```{r}
HOT <- createHOT(
  data = CBC, # data frame
  id = 1, # column index of the id
  none = 21, # column index of none alternative
  prod = 3, # number of alternatives in validation task (excluding none)
  prod.levels = list(c(4, 9, 19), c(8, 12, 17), c(5, 10, 17)), # column index of the attribute levels for each alternative
  coding = c(0, 0, 0), # how the attributes were coded
  method = "CBC", # method
  choice = 22 # column index of choice
)
```

Let us take a glimpse at the output, which shows the total raw utilities for each of the alternatives 
included in the validation/holdout task.

```{r}
head(HOT)
```

In the next step, we would like to see how well our model (from which we took the raw utilities) predict 
the actual choices in the validation/holdout task. First, we will test the <code>hitrate()</code> function. 
We specify the <code>data</code>, the column names of the alternatives (<code>opts</code>; remember 
there are three alternatives + the *no-buy* alternative), and finally the actual choice.

```{r}
hitrate(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice # column name of choice
)
```

Let us also check the magnitude of the mean absolute error by running the <code>mae()</code> function.

```{r}
mae(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice # column name of choice
)
```
To cover also one example on how to use the metrics of the confusion matrix, we could test whether our model overestimates the purchase behavior (our model predicts a *buy* although participant opts for a *no* buy) or underestimates it (i.e., model predicts *no-buy* but participant opts for a *buy*). We will test the accuracy of the model by running the <code>accuracy()</code> function.

```{r}
accuracy(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  none = None # column name of none alternative
)
```

Finally, let us test, how many participants would at least buy one of the three products, assuming that this is one potential assortment we would like to offer to our consumers. We will use the <code>reach()</code> function and use the threshold approach. To specify the bundles we are offering we use the <code>opts</code> argument in our function.

```{r}
reach(
  data = HOT, # data frame
  opts = c(Option_1:Option_3), # products that should be considered
  none = None # column name of none alternative
)
```

### Example II - CBC with linear coding

In a second example, we again use a *CBC*, however, this time we show how to use <code>validateHOT</code> if one of the variables are linear coded. All other examples are provided in the accompanied vignette.

We are using the data frame <code>CBC_lin</code> which is also provided by the <code>validateHOT</code> package. We first load the data frame.

```{r}
data("CBC_lin")
```

Next, we create the validation/holdout task to evaluate it in the next step. We use a validation/holdout task with three alternatives plus the *no-buy* alternative, just as we did in the previous example. The only difference to the previous example is that for our model estimation, the third attribute (<code>Att3_Lin</code>) was coded as linear.

Again, we first define <code>data</code>. The <code>id</code> is saved in the first column. The utilities for the <code>none</code> parameter are stored in the 15th column, and we again set <code>prod</code> to 3 (**excluding** the *no-buy* alternative). Next, we define the <code>prod.levels</code> for each alternative. For example, we can see that the second see that the second alternative is composed of <code>Att1_Lev5</code>, <code>Att2_Lev4</code>, and <code>60</code> which is the value that should be interpolated. We can see that this value is the one that needs to be interpolated if we take a closer look at the <code>coding</code> argument. This tells us that the first two attributes (<code>Att1_Lev5</code>, <code>Att2_Lev4</code> in the case of alternative 2) are part-worth coded (<code>0</code>) while the third attribute is linear coded (<code>1</code>).

To interpolate the value, we have to provide <code>validateHOT</code> the <code>interpolate.levels</code>. These **need** to be the same as provided to Sawtooth as levels. Moreover, it is important that the value that should be interpolated needs to lie within the lower and upper bound of <code>interpolate.levels</code>. In our case, we had 7 levels that range from 10 to 70.

Next, we define the column index of the linear coded variable (<code>lin.p</code>) and specify the <code>coding</code> we talked about above. Again, we are running a *CBC* specified by the <code>method</code> argument. This time, we would like to keep some of the variables in the data frame, which we specify in <code>varskeep</code>. We only keep one further variable, however, you can specify as many as you want. This could be relevant if you would like to display results per group. Finally, we just tell <code>validateHOT</code> the column index of the final choice (<code>choice</code>) and we are all set.

```{r}
CBC <- createHOT(
  data = CBC_lin, # data frame
  id = 1, # column index of the id
  none = 15, # column index of none alternative
  prod = 3, # number of alternatives in validation task (excluding none)
  prod.levels = list(c(4, 9, 60), c(8, 12, 40), c(5, 10, 45)), # column index of the attribute levels for each alternative (for linear coded we specify value to be interpolated)
  interpolate.levels = list(c(10, 20, 30, 40, 50, 60, 70)), # actual values for the levels that should be interpolated
  lin.p = 14, # column index of the linear coded variable
  coding = c(0, 0, 1), # coding of the 3 attributes; 0 = part-worth, 1 = linear, 2 = piecewise
  method = "CBC", # method
  varskeep = 17, # column index of variables that should be kept in the dataframe
  choice = 16 # column index of choice
)
```

The next steps are the same as above. However, let us take a look at some examples in which we display the results per group. Let us again begin with the <code>hitrate()</code> function. To do so, we specify the column name of the grouping variable in the <code>group</code> argument.

```{r}
hitrate(
  data = CBC, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  group = Group # column name of Grouping variable
)
```

In this case, the group assignment is stored in form of integers. However, the output is the same if it is a factor or if it is labelled data. To proof this we just quickly change <code>group</code> into a factor by using the <code>factor()</code> function provided by R Core Team (2023).

```{r}
CBC$Group <- base::factor(
  CBC$Group,
  levels = c(1:3),
  labels = paste0("Group_", c(1:3))
)
```

Afterward, we display the *mean hit probability* by running the <code>mhp()</code> function.

```{r}
mhp(
  data = CBC, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  group = Group # column name of Grouping variable
)
```

For more examples, please see the accompanied vignette.

## References

Burger, Scott V. 2018. <em>Introduction to Machine Learning with R: Rigorous Mathematical Analysis</em>. O'Reilly.

Chrzan, K., & Orme, B. K. (2019). <em>Applied MaxDiff: A Practitioner’s Guide to Best-Worst Scaling</em> Provo, UT: Sawtooth Software.

Ding, Min, John R. Hauser, Songting Dong, Daria Dzyabura, Zhilin Yang, SU Chenting, and Steven P. Gaskin. 2011. “Unstructured Direct Elicitation of Decision Rules.” <em>Journal of Marketing Research 48</em>(1): 116-27. [https://doi.org/10.1509/jmkr.48.1.116](https://doi.org/10.1509/jmkr.48.1.116)

Drost, Hajk-Georg. 2018. “Philentropy: Information Theory and Distance Quantification with R” <em>Journal of Open Source Software 3</em>(26), 765, [https://joss.theoj.org/papers/10.21105/joss.00765](https://joss.theoj.org/papers/10.21105/joss.00765).

Hamner, Ben, and Michael Frasco. 2018. “Metrics: Evaluation Metrics for Machine Learning.” [https://CRAN.R-project.org/package=Metrics](https://CRAN.R-project.org/package=Metrics).

Miaoulis, G., Parsons, H., & Free, V. (1990). Turf: A New Planning Approach for Product Line Extensions. <em>Marketing Research 2</em> (1): 28-40.

Orme, Bryan K. 2015. “Including Holdout Choice Tasks in Conjoint Studies.” [https://sawtoothsoftware.com/resources/technical-papers/including-holdout-choice-tasks-in-conjoint-studies](https://sawtoothsoftware.com/resources/technical-papers/including-holdout-choice-tasks-in-conjoint-studies).

Orme, B. K. (2020). <em>Getting Started with Conjoint Analysis: Strategies for Product Design and Pricing Research</em>. 4th edition. Manhattan Beach, CA: Research Publishers LLC.

R Core Team. 2023. “R: A Language and Environment for Statistical Computing.” [https://www.R-project.org/](https://www.R-project.org/).

Rao, Vithala R. 2014. <em>Applied Conjoint Analysis.</em> Springer Berlin Heidelberg. [https://doi.org/10.1007/978-3-540-87753-0](https://doi.org/10.1007/978-3-540-87753-0).

Yang, Liu (Cathy), Olivier Toubia, and Martijn G. de Jong. 2018. “Attention, Information Processing, and Choice in Incentive-Aligned Choice Experiments.” <em>Journal of Marketing Research 55</em>(6): 783–800. [https://doi.org/10.1177/0022243718817004](https://doi.org/10.1177/0022243718817004).