### 2023-11-11 README.Rmd
updated package - 231111 including markdown

---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# validateHOT 🎯

<!-- badges: start -->

[![R-CMD-check](https://github.com/JoshSchramm94/validateHOT/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JoshSchramm94/validateHOT/actions/workflows/R-CMD-check.yaml)

[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)

<!-- badges: end -->

The goal of `validateHOT` is to validate the results of your validation
task (also known as holdout task) as well as a tool for simulating
markets with the results of your MaxDiff, CBC, and ACBC. This package is
especially relevant for the [Sawtooth
Software](https://sawtoothsoftware.com/) community who would like to
report their analysis in *R* for open science purposes. However, it also
works perfectly with the great package `ChoiceModelR`(Sermas, 2022). Of
course, this package is also useful for practitioners, who would like to
run the analyses in an open source software. The ultimate goal of these
preference measurement techniques is to predict future behavior (Green &
Srinivasan, 1990). Therefore, it is essential for both academics and
practitioners to ensure that the collected data is valid and can also
predict outside tasks that were not included in the estimation of the
utility scores.

Although commercial studies often do not include a validation/holdout
task (Yang et al., 2018), it is highly recommended to do so (Orme, 2015;
Rao, 2014). This validation/holdout task does not only help to check
whether everything went right during data collection but also to
determine the final model. `validatHOT` provides some of the relevant
metrics to test the performance of the data in predicting a
validation/holdout task.

Moreover, another key goal of preference measurement techniques is to
simulate or predict markets (Gilbride, Lenk, & Brazell, 2008).
`validateHOT` provides key functions for an easy implementation to
simulate markets for both (A)CBC and MaxDiff.

> `validateHOT` was developed to work with Sawtooth Software and
> `ChoiceModelR`. Please be cautious about using it with different
> platforms (especially for linear and piecewise-coded variables).

👉🏾 <u>**What you need to provide**</u>: <br> After collecting your data,
and running your initial Hierarchical Bayes models, you can turn to
`validateHOT` and test how good your model predicts choices in the
validation/holdout task. Therefore, you only have to read in your raw
utility scores as well as the actual choice of your validation/ holdout
task. You can use the `merge()` function provided by the *R* base
package to do so (R Core Team, 2023). Afterward, you can read in your
data file and enjoy `validateHOT`. We provide a short tutorial in the
markdown, for a more detailed tutorial, please see the vignette that
comes along with `validateHOT`.

👈🏾 <u>**What you get**</u>:<br> At the moment, `validateHOT` provides
functions for 4 key areas:

<ul>

<li>validation metrics mainly reported in preference measurement
studies</li>

<li>metrics that are usually reported in machine learning (i.e.,
confusion matrix)</li>

<li>simulation methods, for example, to determine optimal product
combinations</li>

<li>converting raw logit utilities in scores that are easier to
interpret</li>

</ul>

For all 4 key areas, the `createHOT()` function is essential. This
function creates the total utilities for each alternative in the
validation/holdout task and also in the simulation method, respectively.
Rao (2014, p. 82) mentions the additive utility model stating that the
total utility of a profile or conjoint model is the sum of its attribute
levels. `createHOT()` will calculate the total utility of each
alternative you specify.

### Classical validation metrics

<ul>

<li><code>hitrate()</code>: creates the Hit Rate (correctly predicted
choices) of your validation/holdout task. The output will contain the
percentage of how many choices were correctly predicted, the chance
level in your validation task ($\frac{1}{alternatives}$) in percentage,
the number of correctly predicted participants' choices as well as the
number of observations.</li>

<li><code>kl()</code>: Kullback-Leibler-Divergence measures the
divergence between the actual choice distribution and the predicted
choice distribution (Ding et al., 2011; Drost, 2018). Output provides
both divergence between predicted from observed and observed from
predicted due to the asymmetry of the Kullback-Leibler divergence. If
you specify an optional <code>group</code> argument the output is split
by groups. Currently, you can choose between \eqn{log} and \eqn{log{_2}}
as logarithm base. The default is set to \eqn{log}.</li>

<li><code>mae()</code>: calculates the mean absolute error, i.e.,
deviation between predicted and stated choice share.</li>

<li><code>medae()</code>: calculates the median absolute error, which is
less affected by outliers compared to the mean absolute error.</li>

<li><code>mhp()</code>: calculates the averaged hit probability of
participant's actual choice in the validation/holdout task.</li>

<li><code>rmse()</code>: provides the root-mean-squared error of
deviation between predicted and stated choice share.</li>

</ul>

All functions can be extended with the `group` argument to get output
split by group(s).

### Confusion Matrix

We also include metrics from machine learning, i.e., the confusion
matrix (e.g., Burger, 2018). For all of the 5 provided functions, you
currently have to include a **none** alternative in your
validation/holdout task. We currently predict, e.g., whether a buy or
no-buy was correctly predicted. Information could be used for
overestimating and underestimating, respectively, of product purchases.
In the following <code>TP</code> stands for true positives,
<code>FP</code> for false positives, <code>TN</code> for true negatives,
and <code>FN</code> for false negatives (Burger, 2018). To translate
this to the `validateHOT` logic, imagine you have a validation/holdout
task with 5 alternatives plus the alternative not to buy any of those
chosen. `validateHOT` now measures whether or not a buy (participant
opts for one of the 5 alternatives) or a no-buy (participant opts for
the no-buy alternative), respectively, is correctly predicted.

Please be aware that for coding, `validateHOT` applies the following
coding of the *buy* and *no-buy* alternatives. Rows refer to the
observed decisions while columns refer to the predicted ones.

|        |     |        |
|--------|:---:|:------:|
|        | Buy | No-buy |
| Buy    | TP  |   FN   |
| No-buy | FP  |   TN   |

<ul>

<li><code>accuracy()</code>: calculates the number of correctly
predicted choices (buy or no-buy); $\frac{TP + TN}{TP + TN + FP + FN}$
(Burger, 2018).</li>

<li><code>f1()</code>: defined as
$\frac{2 * precision * recall}{precision + recall}$ or stated
differently by Burger (2018) $\frac{2TP}{2TP + FP + FN}$.</li>

<li><code>precision()</code>: defined as $\frac{TP}{TP + FP}$ (Burger,
2018).</li>

<li><code>recall()</code>: defined as $\frac{TP}{TP + FN}$ (Burger,
2018).</li>

<li><code>specificity()</code>: defined as $\frac{TN}{TN + FP}$ (Burger,
2018).</li>

</ul>

Again, all functions can be extended with the `group` argument to get
output split by group(s).

### Simulation Methods

<ul>

<li><code>turf()</code>: <b>T</b>(otal) <b>U</b>(nduplicated)
<b>R</b>(each) and <b>F</b>(requency) is a "product line extension
model" (Miaoulis et al., 1990, p. 29) that helps to find the perfect
product bundle based on the reach (e.g., how many participants consider
buying at least one product of that assortment) and the frequency (how
many products are on average a purchase option). <code>turf()</code>
currently provides both the *threshold* approach (<code>approach =
'thres'</code>; all products that exceed a threshold are considered,
e.g., as purchase option; Chrzan & Orme, 2019, p. 112) and the *first
choice* approach (<code>approach = 'thres'</code>; only product with
highest utility is considered as purchase option; Chrzan & Orme, 2019,
p. 111).</li>

<li><code>freqassort()</code>: Similar to <code>turf()</code>,
<code>freqassort()</code> will give you the averaged frequency, how many
products the participants will choose from your in the function
determined potential assortment. Again, you have to define a
<code>none</code> alternative, because <code>freqassort()</code> uses
the <i>threshold</i> approach, meaning if the utility of one product is
above the utility of <code>none</code>, it is marked as potential
purchase option. While <code>turf()</code> calculates the reach and
frequency for <b>all</b> combinations, you specify the combination you
are interested in <code>freqassort()</code>.</li>

<li><code>reach()</code>: Similar to <code>turf()</code>,
<code>reach()</code> will give you the averaged percentage of how many
participants you can reach (at least one of the products resemble a
purchase option) with your in the function determined potential
assortment. <code>reach()</code> also uses the <i>threshold</i> approach
(see above). While <code>turf()</code> calculates the reach and
frequency for <b>all</b> combinations, you specify the combination you
are interested in <code>reach()</code>.</li>

<li><code>marksim()</code>: runs market simulations (either share of
preference or first choice rule), including the standard error, as well
as the lower and upper confidence interval, which is calculated
according to the $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$ (Orme, 2020, p.
94).</li>

</ul>

> <b>Important</b>: For both <code>reach()</code> and
> <code>freqassort()</code>, <code>none</code> does not necessarily have
> to be the no-buy alternative but can be another alternative that
> should be exceeded.

### Reestimating raw utilities

<code>validateHOT</code> also provides 3 functions to better interpret
the scores of both (A)CBC and MaxDiff. Currently, 3 functions are
provided, namely:

<ul>

<li><code>att_imp()</code>: converts the raw utilities of either an ACBC
or CBC into importance scores for each attribute (see, Orme, 2020, pp.
79-81)</li>

<li>

<code>prob_scores()</code>converts the raw utilities of a MaxDiff to
choice probabilities by applying the following procedures:

<ul>

<li>For unanchored MaxDiff: First, the scores are zero-centered,
afterwards they are transformed by the following formula
$\frac{exp^{U_i}}{(exp^{U_i} + a - 1)}$ (Chrzan & Orme, 2019, p. 56),
where $U_i$ is the raw utility of item *i* and <code>a</code> is the
number of items shown simultaneously per MaxDiff task. Finally, the
scores are normalized.</li>

<li>For anchored MaxDiff:
$\frac{exp^{U_i}}{(exp^{U_i} + a - 1)} * 100 * \frac{1}{a}$ (Chrzan &
Orme, 2019, p. 59).</li>

</ul>

</li>

<li><code>zc_diffs()</code>: rescales the raw logit utilities to make
them comparable across participants (Sawtooth Software Inc, 2023, p.
343).</li>

</ul>

### Data Frames provided by <code>validateHOT</code>

`validaeHOT` provides 5 data sets that come with the package. These data
sets should help to better understand the way the data needs to be
defined, especially for the `createHOT()` function.

<ul>

<li><code>ACBC</code>: Example data set with raw utilities of an ACBC
study conducted in Sawtooth. Price was linear-coded while the other
attributes were coded as part-worths.</li>

<li><code>ACBC_interpolate</code>: Example data set with raw utilities
of an ACBC study conducted in Sawtooth. Price was piecewise-coded,
another attribute was linear-coded while the other attributes were coded
as part-worths.</li>

<li><code>CBC</code>: Example data set with raw utilities of a CBC study
conducted in Sawtooth. All attributes were coded as part-worth.</li>

<li><code>CBC_lin</code>: Example data set with raw utilities of a CBC
study conducted in Sawtooth. One attribute was linear coded while the
other attributes are part-worth coded.</li>

<li><code>MaxDiff</code>: Example data set with raw utilities of a
MaxDiff study conducted in Sawtooth.</li>

</ul>

## Why `validateHOT`

We are teaching a preference measurement seminar for students. Often
these students did not have any prior experience (or only sparsely) with
*R*. One of the chapters in this class is about model validating the
results obtained and we teach this, of course, in *R*. We want to put a
strong emphasis on open science and providing tools to run the analyses
and provide the code afterwards is essential.

Of course, there are other great packages which are faster in running
(i.e., `Metrics` by Hamner & Frasco, 2018), however, these packages need
some more data wrangling in order to use the appropriate functions with
the raw utilities, which might be a burden or barrier for one or the
other.

Moreover, as Yang et al. (2018) report, commercial studies often do not
use any validation task. Again, the missing experience in *R* could be
one explanation. Since these functions are not always implemented in
other software, this might be one reason why they do not include one
simply because they do not know how to use it correctly. Having a
package to evaluate the validation/holdout task can also be beneficial
from this perspective.

## Installation

You can install the development version of `validateHOT` from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("JoshSchramm94/validateHOT")
```

## Example

First, we load the package.

```{r}
library("validateHOT")
```

### Example I - CBC

Since *CBC's* are applied more commonly compared to *ACBC* and
*MaxDiff*, we will provide an example with a *CBC*. Let us begin with a
*CBC* where all of the attributes are part-worth coded, so we need no
interpolation. Let us load the `CBC` data frame for this example.

```{r}
data("CBC")
```

The data frame has a total of `r nrow(CBC)` participants and
`r ncol(CBC)` columns.

Now imagine you included a validation/holdout task with 3 alternatives
plus a no-buy alternative. We specify the `data` argument and as well as
the `id`. Since we also have a *no-buy* alternative in our validation
task, we next specify the `none` argument, otherwise we would have left
it empty. Afterwards, we define each alternative with the argument
`prod.levels`. If we look back at the data frame, we can see that the
first alternative in the holdout task (`c(4, 9, 19)`) is composed of the
following attribute levels `Att1_Lev1`, `Att2_Lev1`, and `Att3_Lev5`.

As mentioned above, all the attributes are part-worth coded, therefore,
we set `coding = c(0, 0, 0)`. Finally, we specify the method, which is
`method = "CBC"` in our case, and define the column index of the actual
participant's choice (`choice`). If you run the code, a data frame
called `HOT` (short form of **H**old**o**ut **t**ask) will be returned
to the global environment.

> ❗ `createHOT()` takes both column names and column indexes. However,
> please be aware, if you include linear coding or piecewise coding, you
> **have** to provide the column indexes for the input in prod.levels.,
> please be aware of this. However, you can easily find out the index by
> either using the `names()` function to see all names of the data
> frame, or if you already know the name of the variable, you can use
> `which()` and `colnames()`, both functions are provided by the base
> package (R Core Team, 2023). For example, if you want to find out the
> column index of `ID`, you could run `which(colnames(CBC) == "ID")` to
> determine the column index of `ID`.

```{r}
HOT <- createHOT(
  data = CBC,
  id = "ID",
  none = 21,
  prod.levels = list(c("Att1_Lev1", "Att2_Lev1", "Att3_Lev5"), c(8, 12, 17), c(5, 10, 17)),
  coding = c(0, 0, 0),
  method = "CBC",
  choice = "HOT"
)
```

Let us take a glimpse at the output, which shows the participants' total
raw utilities for each of the 3 alternatives that were included in the
validation/holdout task.

```{r}
head(HOT)
```

In the next step, we would like to see how well our model (from which we
took the raw utilities) predicts the actual choices in the
validation/holdout task. First, we will run the `hitrate()` function. We
specify the `data`, the column names of the alternatives (`opts`;
remember there are 3 alternatives + the *no-buy* alternative), and
finally the actual choice (`choice`).

```{r}
hitrate(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice # column name of choice
)
```

Next, we look at the magnitude of the mean absolute error by running the
`mae()` function. The arguments are the same as for the `hitrate()`
function.

```{r}
mae(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice # column name of choice
)
```

To cover also one example on how to use the metrics of the confusion
matrix, we could test whether our model overestimates the purchase
behavior (our model predicts a *buy* although participant opts for a
*no-buy*) or underestimates it (i.e., model predicts *no-buy* but
participant opts for a *buy*). We will test the accuracy of the model by
running the `accuracy()` function.

```{r}
accuracy(
  data = HOT, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  none = None # column name of none alternative
)
```

Finally, let us test, how many participants would at least buy one of
the 3 products, assuming that this is one potential assortment we would
like to offer to our consumers. We will use the `reach()` function. To
specify the bundles we are offering we use the `opts` argument in our
function.

```{r}
reach(
  data = HOT, # data frame
  opts = c(Option_1:Option_3), # products that should be considered
  none = None # column name of none alternative
)
```

### Example II - CBC with linear coding

In a second example, we again use a *CBC*, however, this time we show
how to use `validateHOT` if one of the variables are linear coded. All
other examples are provided in the accompanied vignette.

We are using the data frame `CBC_lin`. Again, we first load the data
frame.

```{r}
data("CBC_lin")
```

Next, we create the validation/holdout task to evaluate it in the next
step. We use a validation/holdout task with 3 alternatives plus the
*no-buy* alternative, just as we did in the previous example. The only
difference to the previous example is that for our model estimation, the
third attribute (`Att3_Lin`) was coded as linear.

Again, we first define data, namely the `id` as well as the `none`
alternative. Next, we define the `prod.levels` for each alternative.
Since we have one linear coded attribute, we have to specify the column
indexes instead of the column names. For example, for the second
alternative which is composed of `Att1_Lev5`, `Att2_Lev4`, and `60`
which is the value that should be interpolated. We tell `createHOT()`
that this value needs to be interpolated by specifying the `coding`
argument. This tells us that the first two attributes (`Att1_Lev5`,
`Att2_Lev4` in the case of alternative 2) are part-worth coded (`0`)
while the third attribute is linear coded (`1`).

To interpolate the value, we have to provide `createHOT()` the
`interpolate.levels`. These **need** to be the same as provided to
Sawtooth or `ChoiceModelR` as levels. Moreover, it is important that the
value that should be interpolated needs to lie within the lower and
upper bound of `interpolate.levels`. In our case, we had 7 levels that
range from 10 to 70.

Next, we define the column index of the linear coded variable (`lin.p`)
and specify the `coding` we talked about above. Again, we are running a
*CBC* specified by the `method` argument. This time, we would like to
keep some of the variables in the data frame, which we specify in
`varskeep`. We only keep one further variable, however, you can specify
as many as you want. This could be relevant if you would like to display
results per group. Finally, we define the actual choice (`choice`) in
the holdout/validation task and we are all set.

```{r}
CBC <- createHOT(
  data = CBC_lin, # data frame
  id = 1,
  none = 15,
  prod.levels = list(c(4, 9, 60), c(8, 12, 40), c(5, 10, 45)),
  coding = c(0, 0, 1),
  interpolate.levels = list(c(10, 20, 30, 40, 50, 60, 70)),
  lin.p = 14,
  method = "CBC",
  varskeep = "Group",
  choice = "HOT"
)
```

The next steps are the same as above. However, let us take a look at
some examples in which we display the results per group. Let us again
begin with the `hitrate()` function. To do so, we specify the column
name of the grouping variable in the `group` argument.

```{r}
hitrate(
  data = CBC, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  group = Group # column name of Grouping variable
)
```

In this case, the group assignment is stored in form of integers.
However, the output is the same if it is a factor or if it is labelled
data. To proof this we just quickly change `group` into a factor by
using the `factor()` function provided by the *R* Core Team (2023).

```{r}
CBC$Group <- base::factor(
  CBC$Group,
  levels = c(1:3),
  labels = paste0("Group_", c(1:3))
)
```

Afterward, we display the *mean hit probability* by running the `mhp()`
function.

```{r}
mhp(
  data = CBC, # data frame
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  group = Group # column name of Grouping variable
)
```

For more examples, please see the accompanied vignette.

## References

Burger, Scott V. 2018. <em>Introduction to Machine Learning with R:
Rigorous Mathematical Analysis</em>. O'Reilly.

Chrzan, K., & Orme, B. K. 2019. <em>Applied MaxDiff: A Practitioner's
Guide to Best-Worst Scaling</em> Provo, UT: Sawtooth Software.

Ding, Min, John R. Hauser, Songting Dong, Daria Dzyabura, Zhilin Yang,
SU Chenting, and Steven P. Gaskin. 2011. "Unstructured Direct
Elicitation of Decision Rules." <em>Journal of Marketing Research
48</em>(1): 116-27. <https://doi.org/10.1509/jmkr.48.1.116>

Drost, Hajk-Georg. 2018. "Philentropy: Information Theory and Distance
Quantification with R" <em>Journal of Open Source Software 3</em>(26),
765, <https://joss.theoj.org/papers/10.21105/joss.00765>.

Gilbride, Timothy J., Lenk, Peter J., and Brazell, Jeff D. 2008. "Market
Share Constraints and the Loss Function in Choice-Based Conjoint
Analysis." <em>Marketing Science
27</em>(6):995-1011.<https://doi.org/10.1287/mksc.1080.0369>

Green, Paul E., and Srinivasan, V. 1990. "Conjoint Analysis in
Marketing: New Developments with Implications for Research and
Practice." <em>Journal of Marketing 54</em>(4):
3-19.<https://doi.org/10.1177/002224299005400402>.

Sawtooth Software Inc. 2023. <em>Lighthouse Studio 9. Version
9.15.2.</em> Sequim, WA.: Sawtooth Software Inc.

Hamner, Ben, and Michael Frasco. 2018. "Metrics: Evaluation Metrics for
Machine Learning." <https://CRAN.R-project.org/package=Metrics>.

Miaoulis, G., Parsons, H., & Free, V. 1990. Turf: A New Planning
Approach for Product Line Extensions. <em>Marketing Research 2</em> (1):
28-40.

Orme, Bryan K. 2015. "Including Holdout Choice Tasks in Conjoint
Studies."
<https://sawtoothsoftware.com/resources/technical-papers/including-holdout-choice-tasks-in-conjoint-studies>.

Orme, B. K. 2020. <em>Getting Started with Conjoint Analysis: Strategies
for Product Design and Pricing Research</em>. 4th edition. Manhattan
Beach, CA: Research Publishers LLC.

R Core Team. 2023. "R: A Language and Environment for Statistical
Computing." <https://www.R-project.org/>.

Rao, Vithala R. 2014. <em>Applied Conjoint Analysis.</em> Springer
Berlin Heidelberg. <https://doi.org/10.1007/978-3-540-87753-0>.

Sermas R (2022). *ChoiceModelR: Choice Modeling in R*. R package version
1.3.0, <https://CRAN.R-project.org/package=ChoiceModelR>.

Yang, Liu (Cathy), Olivier Toubia, and Martijn G. de Jong. 2018.
"Attention, Information Processing, and Choice in Incentive-Aligned
Choice Experiments." <em>Journal of Marketing Research 55</em>(6):
783--800. <https://doi.org/10.1177/0022243718817004>.